---
title: "Capstone Project"
author: "Eric Kao"
date: "March 26, 2017"
output: html_document
---



## Student Alcohol Consumption

**Student Alcohol Consumption** dataset is what I decided for my springboard capstone project. You can find the original dataset at [here](https://www.kaggle.com/uciml/student-alcohol-consumption)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(plyr)
#preload all the dataframe
math <-read.csv('student-mat.csv')##math class grade
portuguese <-read.csv('student-por.csv')##portugese

join_set=merge(math,portuguese,by=c("school","sex","age","address","famsize","Pstatus", "Medu","Fedu","Mjob","Fjob","reason","nursery","internet", "guardian","guardian","traveltime","studytime","failures", "schoolsup","famsup","activities","higher","romantic", "famrel","freetime","goout","Dalc","Walc","health","absences"))



```


Let's take a first look of the dataset.

```{r}

str(join_set)

```



## Including Plots

###Data wrangling

I create grade_mean_portuguese and grade_mean_math to record the average of the student and create another column total_mean to record the average of grade_mean_math and grade_mean_portuguese.

```{r}
library(ggplot2)
join_set<-join_set%>%mutate(grade_mean_math=(G1.x+G2.x+G3.x)/3)%>%mutate(grade_mean_portuguese=(G1.y+G2.y+G3.y)/3)%>%mutate(total_mean = ((grade_mean_math+grade_mean_portuguese)/2) )%>%mutate(schoolsup=as.factor(schoolsup),famsup=as.factor(famsup),paid.x=as.factor(paid.x),paid.y=as.factor(paid.y),activities=as.factor(activities),nursery=as.factor(nursery),higher=as.factor(higher),internet=as.factor(internet),romantic=as.factor(romantic)) 
```

Next, I change Dalc and Walc into text in order to read the value easier. 
```{r}
DalcV <- mapvalues(join_set$Dalc, from = 1:5, to = c("Very Low", "Low", "Medium", "High", "Very High"))
WalcV <- mapvalues(join_set$Walc, from = 1:5, to = c("Very Low", "Low", "Medium", "High", "Very High"))

```

When I do this I notice the there is no "High" and "Very high" in the data set. 


Let's see the gender structure of this sample.
```{r}
ggplot(join_set, aes(x=sex,col=sex,fill=sex))+geom_bar(position= "dodge")+ggtitle("Gender composition")
```
In this sample, there are 50 females and 35 males. 


###What's the consumption distribution for both group
```{r}
ggplot(join_set, aes(x=sex,  fill=DalcV)) +
geom_bar(stat="count")+ggtitle(" Workday alcohol consumption distribution based on genders")
```

```{r}
ggplot(join_set, aes(x=sex,  fill=WalcV)) +geom_bar(stat="count")+ggtitle(" Weekend alcohol consumption distribution based on genders")
```

This graph also show that there is a large portion of both group consume very low alcohol. 


## Alcohol consumption and Student's grade relationship



```{r}
Dalc_total_mean<-ggplot(join_set,aes(x=reorder(DalcV,total_mean, order = FALSE),y= total_mean))
Dalc_total_mean+geom_boxplot()+labs(x="Workday Alcohol Consumption")+ggtitle(" Workday alcohol consumption boxplot and average grade ")
```


From this plot, the group that consume "Medium" alcohol have a lower median then other 2 groups. The "Median" group also have a lower average than other 2 groups.



```{r}
Walc_total_mean = ggplot(join_set,aes(x=reorder(Walc,total_mean, order = FALSE),y= total_mean))

Walc_total_mean+geom_boxplot(aes(x=reorder(WalcV,total_mean, order = FALSE)))+labs(x="Workday Alcohol Consumption")+ggtitle(" Weekend alcohol consumption boxplot and average grade ")
```
The "High" group have the lowest total_mean and the "Medium" group have the second lowest total_mean. Surprisingly, the "Very High" have the second highest total_mean. 

```{r}
ggplot(join_set,aes(x=as.factor(WalcV),y= as.factor(DalcV), col = as.factor((DalcV))))+geom_jitter()

```



##Health and Student grade relationship

```{r}
Health_total_mean =  ggplot(join_set,aes(x=as.factor(health),y= total_mean))
Health_total_mean + geom_boxplot()+labs(x= "Health status")
```
According to the plot, the group of student who has the worst health condition perform actually better than the other group. The healthiest group has the widest distribution in the total_mean. 




##Health and alcohol consumption relationship

```{r}
ggplot(join_set, aes(x= as.factor(Dalc), y = as.factor(health) ))+geom_jitter(aes(col= sex))+ggtitle("Student health distribution dotplot based on Workday alcohol consumption")+labs(x="Workday Alcohol Consumption")+facet_grid(.~health)
```
From this plot, the healthiest group does not related to the least alcohol consumption. On contrary, the least healthy group have no student is in "Medium" alcohol consumption group. 

```{r}
ggplot(join_set, aes(x= as.factor(Walc), y = as.factor(health) ))+geom_jitter(aes(col= sex))+ggtitle("Student health distribution dotplot based on weekend alcohol consumption")+labs(x="Weekend Alcohol Consumption")+facet_grid(.~health)
```

The least healthy group have no one falls in the "Very high" and "high" groups; however in the "low" group of health, there are 5 student in the "Very high" and "high" groups.

```{r}
Dalc_total_mean+geom_jitter(aes(col= sex))+ggtitle("Student total_mean distribution dotplot based on Workday alcohol consumption")+labs(x="Workday Alcohol Consumption")
```

```{r}
ggplot(join_set,aes(x=sex,col=sex))+geom_bar(position="dodge",stat="count",aes(fill=sex))+facet_grid(. ~ reorder(Dalc,total_mean, order = FALSE))+ggtitle("Student headcount of each Workday alcohol consumption group")
```
Most of female students are in the "Very Low" group and there is no female students in "Medium" group. 







The distribution of gender between schools are the same. 


```{r}

#why this is not woring in here
base1_class_sex<-ggplot(join_set,aes(x=school, col=sex))
base1_class_sex+geom_bar(aes(fill=sex),stat="count",position = "fill")+facet_grid(.~Walc)+ggtitle("Gender Proportion of each group")

```
Gender proportion of each daily consumption group. Male tends to drink more alcohol than female. 

#Data Modeling

set up the test set and training set.

```{r}
library(caTools)
set.seed(144)
split = sample.split(join_set$total_mean, SplitRatio = 0.65)
train = subset(join_set,split==TRUE)
test = subset(join_set,split==FALSE)
```


### Do the linear regression first
```{r}

Train.lm = lm(total_mean ~.-G1.x-G2.x-G3.x-G1.y-G2.y-G3.y-grade_mean_math-grade_mean_portuguese,data=train)
Train.lm.improve = lm(total_mean ~Mjob+reason+nursery+romantic+absences+paid.x,data=train)

summary(Train.lm)
```


```{r}
Traim.lm.predict= predict(Train.lm, newdata = test)

SSE_pre= sum((test$total_mean-Traim.lm.predict)^2)
SST_pre= sum((test$total_mean-mean(train$total_mean))^2)
R_sqare= 1-(SSE_pre/SST_pre)
SSE_pre
SST_pre
R_sqare
```



##Random Forest
```{r}
library(randomForest)
train$Dalc = as.factor(train$Dalc)
train$Walc = as.factor(train$Walc)
test$Dalc = as.factor(test$Dalc)
test$Walc = as.factor(test$Walc)
TrainrForest = randomForest( total_mean ~ . -G1.x-G2.x-G3.x-G1.y-G2.y-G3.y-grade_mean_math-grade_mean_portuguese, data = train, nodesize= 25, ntree=200)
PredictForest = predict(TrainrForest, newdata = test)
SSE_pre= sum((test$total_mean-PredictForest)^2)
SST_pre= sum((test$total_mean-mean(train$total_mean))^2)
R_sqare= 1-(SSE_pre/SST_pre)

SSE_pre
SST_pre
R_sqare
```

##Decision tree
```{r}
library(rpart)
library(rpart.plot)
TrainTree = rpart(total_mean~. -G1.x-G2.x-G3.x-G1.y-G2.y-G3.y-grade_mean_math-grade_mean_portuguese, data = train,  control = rpart.control(minbucket = 25))
prp(TrainTree)

```
Decision tree R-square
```{r}
tree.pred = predict(TrainTree,newdata = test)
tree.sse = sum((tree.pred-test$total_mean))
tree.sst= sum((test$total_mean-mean(train$total_mean))^2)
tree.Rsquare = R_sqare= 1-(tree.sse/tree.sst)
tree.sse
tree.sst
tree.Rsquare
```




