---
title: "Capstone Project"
author: "Eric Kao"
date: "March 26, 2017"
output: html_document
---



## Student Alcohol Consumption

**Student Alcohol Consumption** dataset is what I decided for my springboard capstone project. You can find the original dataset at [here](https://www.kaggle.com/uciml/student-alcohol-consumption)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(plyr)
#preload all the dataframe
math <-read.csv('student-mat.csv')##math class grade
portuguese <-read.csv('student-por.csv')##portugese

join_set=merge(math,portuguese,by=c("school","sex","age","address","famsize","Pstatus", "Medu","Fedu","Mjob","Fjob","reason","nursery","internet", "guardian","guardian","traveltime","studytime","failures", "schoolsup","famsup","activities","higher","romantic", "famrel","freetime","goout","Dalc","Walc","health","absences"))



```


Let's take a first look of the dataset.

```{r}

str(join_set)

```



## Including Plots

###Data wrangling

I create grade_mean_portuguese and grade_mean_math to record the average of the student and create another column total_mean to record the average of grade_mean_math and grade_mean_portuguese.

```{r}
library(ggplot2)
join_set<-join_set%>%mutate(grade_mean_math=(G1.x+G2.x+G3.x)/3)%>%mutate(grade_mean_portuguese=(G1.y+G2.y+G3.y)/3)%>%mutate(total_mean = ((grade_mean_math+grade_mean_portuguese)/2) )%>%mutate(schoolsup=as.factor(schoolsup),famsup=as.factor(famsup),paid.x=as.factor(paid.x),paid.y=as.factor(paid.y),activities=as.factor(activities),nursery=as.factor(nursery),higher=as.factor(higher),internet=as.factor(internet),romantic=as.factor(romantic)) 
```

Next, I change Dalc and Walc into text in order to read the value easier. 
```{r}
DalcV <- mapvalues(join_set$Dalc, from = 1:5, to = c("Very Low", "Low", "Medium", "High", "Very High"))
WalcV <- mapvalues(join_set$Walc, from = 1:5, to = c("Very Low", "Low", "Medium", "High", "Very High"))

```

When I do this I notice the there is no "High" and "Very high" in the data set. 


Let's see the gender structure of this sample.


```{r}
ggplot(join_set, aes(x=sex,col=sex,fill=sex))+geom_bar(position= "dodge")+ggtitle("Gender composition")
```


In this sample, there are 50 females and 35 males. 


###What's the consumption distribution for both group
```{r}
ggplot(join_set, aes(x=sex,  fill=DalcV)) +
geom_bar(stat="count")+ggtitle(" Workday alcohol consumption distribution based on genders")
```

```{r}
ggplot(join_set, aes(x=sex,  fill=WalcV)) +geom_bar(stat="count")+ggtitle(" Weekend alcohol consumption distribution based on genders")
```



This graph also show that there is a large portion of both group consume very low alcohol. 


## Alcohol consumption and Student's grade relationship



```{r}
Dalc_total_mean<-ggplot(join_set,aes(x=reorder(DalcV,total_mean, order = FALSE),y= total_mean))
Dalc_total_mean+geom_boxplot()+labs(x="Workday Alcohol Consumption")+ggtitle(" Workday alcohol consumption boxplot and average grade ")
```




From this plot, the group that consume "Medium" alcohol have a lower median then other 2 groups. The "Median" group also have a lower average than other 2 groups.



```{r}
Walc_total_mean = ggplot(join_set,aes(x=reorder(Walc,total_mean, order = FALSE),y= total_mean))

Walc_total_mean+geom_boxplot(aes(x=reorder(WalcV,total_mean, order = FALSE)))+labs(x="Workday Alcohol Consumption")+ggtitle(" Weekend alcohol consumption boxplot and average grade ")
```



The "High" group have the lowest total_mean and the "Medium" group have the second lowest total_mean. Surprisingly, the "Very High" have the second highest total_mean. 

```{r}
ggplot(join_set,aes(x=as.factor(WalcV),y= as.factor(DalcV), col = as.factor((DalcV))))+geom_jitter()

```



##Health and Student grade relationship

```{r}
Health_total_mean =  ggplot(join_set,aes(x=as.factor(health),y= total_mean))
Health_total_mean + geom_boxplot()+labs(x= "Health status")
```




According to the plot, the group of student who has the worst health condition perform actually better than the other group. The healthiest group has the widest distribution in the total_mean. 




##Health and alcohol consumption relationship

```{r}
ggplot(join_set, aes(x= as.factor(Dalc), y = as.factor(health) ))+geom_jitter(aes(col= sex))+ggtitle("Student health distribution dotplot based on Workday alcohol consumption")+labs(x="Workday Alcohol Consumption")+facet_grid(.~health)
```



From this plot, the healthiest group does not related to the least alcohol consumption. On contrary, the least healthy group have no student is in "Medium" alcohol consumption group. 

```{r}
ggplot(join_set, aes(x= as.factor(Walc), y = as.factor(health) ))+geom_jitter(aes(col= sex))+ggtitle("Student health distribution dotplot based on weekend alcohol consumption")+labs(x="Weekend Alcohol Consumption")+facet_grid(.~health)
```



The least healthy group have no one falls in the "Very high" and "high" groups; however in the "low" group of health, there are 5 student in the "Very high" and "high" groups.






#Data Modeling

Set up the test set and training set.

```{r}
library(caTools)
set.seed(144)
split = sample.split(join_set$total_mean, SplitRatio = 0.65)
train = subset(join_set,split==TRUE)
test = subset(join_set,split==FALSE)
```


### Linear Regression
```{r}
Train.lm = lm(total_mean ~.-G1.x-G2.x-G3.x-G1.y-G2.y-G3.y-grade_mean_math-grade_mean_portuguese,data=train)
summary(Train.lm)
Train.lm.improve = lm(total_mean ~Mjob+reason+nursery+romantic+absences+paid.x,data=train)
summary(Train.lm.improve)
Train.lm.improve2 = lm(total_mean ~Mjob+nursery+romantic,data=train)
summary(Train.lm.improve2)
```


```{r}
#Traim.lm.predict= predict(Train.lm, newdata = test)
#Traim.lm.predict= predict(Train.lm.improve, newdata = test)
Traim.lm.predict= predict(Train.lm.improve2, newdata = test)#improved model
SSE_pre= sum((test$total_mean-Traim.lm.predict)^2)
SST_pre= sum((test$total_mean-mean(train$total_mean))^2)
R_sqare= 1-(SSE_pre/SST_pre)
MSRE= sqrt(SSE_pre/nrow(test))
SSE_pre
SST_pre
R_sqare
MSRE
```
Before improving the model, the r-square is -5.49. RSME is 7.66.
After improving, the r-square is -0.3074805. RSME is 3.440894.



##Random Forest
```{r}
library(randomForest)
train$Dalc = as.factor(train$Dalc)
train$Walc = as.factor(train$Walc)
test$Dalc = as.factor(test$Dalc)
test$Walc = as.factor(test$Walc)
TrainrForest = randomForest( total_mean ~ . -G1.x-G2.x-G3.x-G1.y-G2.y-G3.y-grade_mean_math-grade_mean_portuguese, data = train, nodesize= 25, ntree=200)
PredictForest = predict(TrainrForest, newdata = test)
rf.SSE_pre= sum((test$total_mean-PredictForest)^2)
rf.SST_pre= sum((test$total_mean-mean(train$total_mean))^2)
rf.R_sqare= 1-(rf.SSE_pre/rf.SST_pre)
rf.MSRE = sqrt(rf.SSE_pre/nrow(test))

rf.SSE_pre
rf.SST_pre
rf.R_sqare
rf.MSRE
```
The r-square is 0.1134692. RSME is 2.833353.

##Decision tree
```{r}
library(rpart)
library(rpart.plot)
TrainTree = rpart(total_mean~. -G1.x-G2.x-G3.x-G1.y-G2.y-G3.y-grade_mean_math-grade_mean_portuguese, data = train)
prp(TrainTree)

```



Decision tree R-square

```{r}
tree.pred = predict(TrainTree,newdata = test)
tree.sse = sum((tree.pred-test$total_mean))
tree.sst= sum((test$total_mean-mean(train$total_mean))^2)
tree.Rsquare= 1-(tree.sse/tree.sst)
tree.RSME = sqrt(tree.sse/nrow(test))
tree.sse
tree.sst
tree.Rsquare
tree.RSME

```
The r-square is 0.9719373. RSME is 0.5041024.


##Conclusion
In the sample distribution, the most of the student do not drink much alcohol, but on the weekend the restriction are slightly loose. In the analysis, it point out that the performance of students does not have a direct relationship with weekend alcohol consumption but if the student consume alcohol in the workdays, it has slightly negative relationship. 

The prediction model shows an interesting conclusion because the student who has family educational support should perform better in the test. On contrary, those students are predicted to have a lower score. I assume the family educational support might kill the student’s interest of learning. Nowaday, the cram school education are thriving in the US and that can possibly put out the student’s interest of learning.

After all, I think the school authority should think about how to ignite students’ passion of learning. No matter how good the education program can be, if the teacher cannot give the student the will of learning, it will be a waste. Parents, especially the mother also plays a vital role in the student’s education. I used to think if one of the parents is teacher, the student can performance better since the student can discuss the school work with the parent whenever they want. In this case, the student has the worst performance when the mother is a teacher. I can imagine a world that the student collapse because one of the parents is a teacher and pressure the student hard. In the end, if the student is not interested in learning, no one will be able to force it. 



