---
title: "Capstone Project"
author: "Eric Kao"
date: "March 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(plyr)
#preload all the dataframe
df1 <-read.csv('student-mat.csv')
df2 <-read.csv('student-por.csv')

d3=merge(df1,df2,by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))

```

## Student Alcohol Consumption

**Student Alcohol Consumption** dataset is what I decided for my springboard capstone project. You can find the original dataset at [here](https://www.kaggle.com/uciml/student-alcohol-consumption)

Let's take a first look of the dataset.

```{}
View(df1)
View(df2)


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(ggplot2)
base1<-ggplot(df1,aes(x=Dalc,y=G1,col=sex))
base1+geom_point()
base1+geom_jitter()
base1+geom_boxplot()+facet_grid(.~Dalc)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
